For submission to the Journal of Integrated Pest Management

### 2021-05-17 ofv_fife_2021.Rmd
Changed the .csl from 'disambiguate-add-names: true' to 'false' and set 'disambiguate-add-year-suffix' to 'true' to make citations add an 'a' or 'b' to the year in inline citations, instead of showing the second author to disambiguate publications in the same year with the same first author. This is a modification of the ESA's .csl file.

---
title: "For submission to the Journal of Integrated Pest Management"
subtitle: "Category: 'Issues'"
output: 
  word_document:
    reference_docx: ../r-word-templates/word-template-jis.docx
# - html_document
bibliography: 'bib/thesis.bib'
csl: 'j_med_entomol.csl'
#always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

@@

*Orchid fleck dichorhavirus* (OFV) infects over fifty plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (Citrus). The sole vectors for dichorhaviruses are flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae), which are also known to spread cileviruses. OFV was found infecting liriopogons *Liriope* spp. and *Ophiopogon* spp.,  as well as *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in the landscape in Leon County, Florida, during the summer of 2020. The presence of OFV was confirmed using OFV-specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98% identity with the known OFV sequences available in the NCBI GenBank. Virus identity was also confirmed via quantitative RT-PCR (RT-qPCR). Additional leaf samples were collected from other possibly infected Asparagaceae from Leon and Alachua counties. Identification of partial genome sequence confirmed the presence of both OFV orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. These strains of OFV are known to infect citrus and cause citrus leprosis disease. Three potential mite vectors were identified from OFV-infected plants using cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* s.l., *B. obovatus*, and *B. confusus*. Florida has various native and introduced plants in the landscape that Brevipalpus spp. feed on, which are potentially susceptible to OFV. In this study, we report three new hosts from the family Asparagaceae from multiple locations. Our data suggested that OFV is widely distributed in Florida and might be a threat for various plant species growing in the southeastern US.

@@

\pagebreak
*Orchid fleck dichorhavirus* (OFV) is the type member for the genus *Dichorhavirus*, family Rhabdoviridae; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the sole vectors for cileviruses and dichorhaviruses [@Maeda1998]. *Brevipalpus californicus* (Banks) sensu lato mites are the only group known to transmit OFV in a persistent propagative manner [@Kondo2003].

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (`r citef("fig_1")`, `r citef("fig_2")`).

@@

### Conclusion
The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence. According to the International Committee on Taxonomy of Viruses (ICTV) classification, OFV consist of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2). Both the orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the US multi-billion dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of Dichorhaviruses (OFV) and Cileviruses (Citrus Leprosis) [@Knorr1968; @Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`). Therefore, it is critical to identify the vector of OFVs in Florida and monitor its spread to determine the risk this virus represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

>>>>

---
title: "For submission to the Journal of Integrated Pest Management"
subtitle: "Category: 'Issues'"
output: 
  word_document:
    reference_docx: ../r-word-templates/word-template-jis.docx
# - html_document
bibliography: 'bib/thesis.bib'
citation_package: natbib
biblio-style: apalike
csl: 'entsoc.csl'
#always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

@@

*Orchid fleck dichorhavirus* (OFV) infects over 50 plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (Citrus). The only known vectors of dichorhaviruses are flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae). OFV was found infecting liriopogons (*Liriope* spp. and *Ophiopogon* spp.),  as well as *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in the landscape in Leon County, Florida, during the summer of 2020. The presence of OFV was confirmed using OFV-specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98% identity with known OFV sequences available in NCBI GenBank. Other leaf samples of the plant family Asparagaceae with ringspots and chlorosis were collected from Leon and Alachua counties and tested with PCR and Sanger sequencing to detect OFV. Identification of partial genome sequence from PCR products confirmed the presence of both OFV orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. These strains of OFV are known to infect citrus and cause citrus leprosis disease. Three potential mite vectors were identified from OFV-infected plants using cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (sensu lato), *B. obovatus*, and *B. confusus*. Florida has various native and introduced plants in the landscape that Brevipalpus spp. feed on, which are potentially susceptible to OFV. In this study, we report three new hosts from the family Asparagaceae from multiple locations. Our data suggested that OFV is widely distributed in Florida and might be a threat for various plant species growing in the southeastern US.

@@

\pagebreak
*Orchid fleck dichorhavirus* (OFV) is the type member for the genus *Dichorhavirus*, family Rhabdoviridae; a bacilliform, nuclear rhabdovirus composed of two segments of single-stranded, negative-sense RNA which infects plants [@Amarasinghe2019; @Dietzgen2014; @Walker2018]. Flat mites from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae) are the sole vectors of dichorhaviruses [@Maeda1998]. *Brevipalpus californicus* (Banks) sensu lato mites are the only group known to transmit OFV in a persistent propagative manner [@Kondo2003].

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns<!--(`r citef("fig_1")`, `r citef("fig_2")`)-->.

@@

### Conclusion
The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence. According to the International Committee on Taxonomy of Viruses (ICTV) classification, OFV consist of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2). Both the orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011a found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the US multi-billion dollar citrus industry. *B. californicus*, as well as *B. yothersi* (Baker), are both known vectors of Dichorhaviruses (OFV) [@Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`). Therefore, it is critical to identify the vector of OFVs in Florida and monitor its spread to determine the risk this virus represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.
